---
title: "practice exam"
author: XIYU Yang
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#rm(list = ls())

library(ggplot2)
library(tibble)
library(tidyr)
library(readr)

library(purrr)
library(dplyr)
library(stringr)
library(forcats)

library(nycflights13)
library(lubridate)
```

#### Make a plot with three facets, one for each airport in the weather data. The x-axis should be the day of the year (1:365) and the y-axis should be the mean temperature recorded on that day, at that airport.

```{r, warning=FALSE}
# Load the dataset
weather_df1 <- weather %>% 
  mutate(day_of_year = yday(time_hour)) %>% 
  group_by(day_of_year) %>% 
  mutate(avg_temp = mean(temp)) %>% 
  select(day_of_year, avg_temp, origin) %>% 
  distinct()

# Create facets from the data
p <-  ggplot(data = weather_df1,
             mapping = aes(x = day_of_year,
                           y = avg_temp,
                           color = origin))

p + geom_point() +
  facet_wrap(~ origin)
```

#### Make a non-tidy matrix of that data where each row is an airport and each column is a day of the year.

```{r}
weather_df2 <- weather_df1 %>% 
  pivot_wider(names_from = day_of_year, values_from =avg_temp)
head(weather_df2)
```

#### For each (airport, day) contruct a tidy data set of the airport's "performance" as the proportion of flights that departed less than an hour late.

```{r, comment=NA, warning=FALSE}
flights_df1 <- flights %>% 
  mutate(day_of_year = yday(time_hour)) %>% 
  group_by(origin, day_of_year) %>% 
  mutate(n1 = n()) %>% 
  filter(dep_delay < 60) %>% 
  mutate(n2 = n()) %>% 
  mutate(performance=n2/n1) %>% 
  group_by(origin, day_of_year) %>% 
  select(origin, day_of_year, performance) %>% 
  distinct()
head(flights_df1)
```

#### Construct a tidy data set to that give weather summaries for each (airport, day). Use the total precipitation, minimum visibility, maximum wind_gust, and average wind_speed.

```{r, warning=FALSE}
weather_df3 <- weather %>% 
  mutate(day_of_year = yday(time_hour)) %>% 
  group_by(origin, day_of_year) %>% 
  summarise(total_precipitation=sum(precip, na.rm=T),
            min_visibiligy=min(visib, na.rm=T),
            max_wind_gust=max(wind_gust, na.rm=T),
            avg_wind_speed=mean(wind_speed, na.rm=T))
head(weather_df3)
```

#### Construct a linear model to predict the performance of each (airport,day) using the weather summaries and a "fixed effect" for each airport. Display the summaries.

```{r}
# Combine the flight and the weather datasets
combined <- flights_df1 %>% 
  left_join(weather_df3, by=c('origin', 'day_of_year')) %>% 
  filter(!is.na(max_wind_gust),
         !is.na(total_precipitation),
         !is.na(min_visibiligy),
         !is.na(avg_wind_speed))

combined$max_wind_gust[which(combined$max_wind_gust == -Inf)] <- 0

# Construct the multiple linear regression model
lm1 <- lm(performance~origin+total_precipitation+min_visibiligy
         +max_wind_gust+avg_wind_speed, data=combined)
summary(lm1)
```

#### Repeat the above, but only for EWR. Obviously, exclude the fixed effect for each airport.

```{r, warning=FALSE, comment=NA}
combined_EWR <- combined %>% filter(origin == 'EWR')
lm2 <- lm(performance~total_precipitation+min_visibiligy
         +max_wind_gust+avg_wind_speed, data=combined_EWR)
summary(lm2)
```





